\LARGE Performance Evaluation of Xenomai 3

### 2015-10-21 paper.tex
improve test descriptions

\section{Performance Evaluation}
There are two objectives for performance evaluations. First, compare the performance of Xenomai with baseline Linux kernel and PREEMPT\_RT patched kernel to decide which configuration has better realtime capability. Second, try to validate whether Xenomai-3 has improved its realtime performance. The details are described in following sub-chapters. 
\subsection{Environment Setup}
Because building and transporting operating systems on embedded devices is complicated. In order for other people to re-do the experiment, it needs proper setup and documentation. We will elaborate on different parts of the experiment setup. 
\subsubsection{Hardware}
For the testing platform, we used the BeagleBone Black. BeagleBone Black is an effective embedded platform with ARM Cortex-A8 processor, capable of running Linux and Xenomai (both 2 and 3), consisting of a single processor, a 512 MB memory, 32kB L1 caches and a 256kB L2 cache.  
\subsubsection{Linux Kernel}
We installed a Debian Linux image\cite{debian} on the beaglebone, with hardware-specific linux kernel from\cite{kernel}. Kernel version 3.14.39 were chosen due to it's compatibility with all kernel configurations.
Four distinct kernel configurations are used for the experiment, as summarized in table below. Kernel are patched using git merge first, and remaining conflicts are resolved manually.

Kernel conﬁguration options are mostly left as default. However, some of the options may have an impact on real-time latencies. These options include all options under CPU Power Management, APM support, and Opportunistic sleep. Additionally, most debug options in Kernel hacking section lead to increased latencies. Those options are turned off before compilation.

\begin{tabular}{|l| p{3cm} |l|}
\hline
Kernel & Patches & Source\\ \hline
mainline   & none & n/a \\ \hline
PREEMPT\_RT & patch-3.14.39-rt37 & \cite{p-rt} \\ \hline
Xenomai 2  & ipipe-core-3.14.39-arm-9 + xenomai v2.6.4 (2015-07-13 HEAD) & \cite{p-ipipe} \cite{git-x2.6} \\ \hline
Xenomai 3  & ipipe-core-3.14.39-arm-9 + xenomai v3.0  & \cite{p-ipipe} \cite{git-x3} \\ \hline
\end{tabular}\\
Table 1: Different kernel configurations for the experiment\\
\subsubsection{Pre-experiment Settings}
We tried to measure the latencies under 2 conditions - IDLE and CPU-STRESSED. In the IDLE experiment, in order to create such test environment, some redundant user processes were eliminated. While in the CPU-STRESSED condition, these process are also killed, and we used the 'stress' command to keep the CPU busy. The command to kill these processes are written in the test script.

@@

\bibliographystyle{unsrt}
\begin{thebibliography}{9}%use this if you have <=9 bib refs
	\bibitem {rtai}{\it https://www.rtai.org/}
	\bibitem {sel4}{\it http://ssrg.nicta.com.au/projects/seL4/}
	\bibitem {linux-rt}{\it http://git.kernel.org/cgit/linux/kernel/git/rt/}
	\bibitem {xenomai}{\it http://xenomai.org/}
	\bibitem {adeos} Karim Yaghmour. Adaptive Domain Environment for Operating Systems.\\
	    \url{http://www.opersys.com/ftp/pub/Adeos/adeos.pdf}
	\bibitem {ChameleonRTOS}{\it http://www.xenomai.org/documentation/slides/Xenomai-OSMB-2007-01.pdf}
	\bibitem {posix-1003-1c}{\it http://standards.ieee.org/findstds/interps/1003-1c-95\_int/}
	\bibitem {xenomai-solo}{\it https://www.osadl.org/Xenomai-SOLO.xenomai-solo.0.html}
	\bibitem {x3-applications}{\it http://xenomai.org/running-applications-with-xenomai-3-x/}
	\bibitem {debian}{\it http://debian.beagleboard.org/images/bone-debian-7.5-2014-05-14-2gb.img.xz}
	\bibitem {kernel}{\it https://github.com/beagleboard/linux}
	\bibitem {p-rt}{\it https://www.kernel.org/pub/linux/kernel/projects/rt/3.14/}
	\bibitem {p-ipipe}{\it http://download.gna.org/adeos/patches/v3.x/arm/ipipe-core-3.14.39-arm-9.patch}
	\bibitem {git-x2.6}{\it http://git.xenomai.org/xenomai-2.6.git/}
	\bibitem {git-x3}{\it http://git.xenomai.org/xenomai-3.git/}
\end{thebibliography}

>>>>

\section{Performance Evaluation}
There are two objectives for performance evaluations. First, comparing the performance of Xenomai with mainline Linux kernel and PREEMPT\_RT to exploit the real-time capability from system view. Second, trying to quantize the performance improvements and potential degradation of Xenomai 3.

\subsubsection{Hardware}
\textit{BeagleBone Black}\cite{bbb} is selected as our major testing hardware platform powered by \textit{TI AM335x} \cite{am335x}, and it is an effective embedded platform, capable of running Linux and Xenomai (both 2 and 3), consisting of a single 1GHz ARM Cortex-A8 processor, a 512 MB memory, 32kB L1 caches and a 256kB L2 cache.

Beaglebone Black is community-supported development platform, and active Linux kernel maintenance enables us to perform further experiments and participate in technical discussions. TI AM335x is used widely in industrial automation, providing an extensive, reliable solution portfolio ranging from robust microcontrollers to ARM based microprocessors.

\subsubsection{Software Installation}
Debian GNU/Linux image 2014-05-14 \cite{debian} was installed on the BeagleBone Black, with hardware-specific Linux kernel tree from official BeagleBone kernel repository \cite{kernel}. We selected kernel version 3.14.39 due to its compatibility with all kernel configurations after our early evaluations.

Four distinct Linux kernel configurations are used for the experiments, as summarized in table below. Kernel are patched using git merge first, and remaining conflicts are resolved manually.

The kernel conﬁguration options are mostly following the default. However, some of the default configuration may have an impact on real-time latencies, such as the ones about CPU Power Management, APM support, and opportunistic sleep. Additionally, most debugging facilities in Kernel hacking section lead to increased latencies. Those problematic options are turned off manually in advance.

\begin{tabular}{|l| p{3cm} |l|}
\hline
Kernel & Patches & Source\\ \hline
mainline   & N/A & N/A \\ \hline
PREEMPT\_RT & patch-3.14.39-rt37 & \cite{p-rt} \\ \hline
Xenomai 2  & ipipe-core-3.14.39-arm-9 + xenomai v2.6.4 (Jul 13, 2015; HEAD) & \cite{p-ipipe} \cite{git-x2.6} \\ \hline
Xenomai 3  & ipipe-core-3.14.39-arm-9 + xenomai v3.0 (Oct 8, 2015) & \cite{p-ipipe} \cite{git-x3} \\ \hline
\end{tabular}\\
Table 1: Different kernel configurations for the experiment\\

\subsubsection{Prerequisites for Experiments}
We tried to measure the latencies under 2 conditions: IDLE and CPU-STRESSED. In the IDLE condition, in order to create such test environment, some redundant user processes were eliminated. While in the CPU-STRESSED condition, these process are also forcely killed, and we used the \textit{stress}\cite{stress} utility to keep the CPU busy. The command to kill these processes are written in the test script.

@@

\bibliographystyle{unsrt}
\begin{thebibliography}{9}%use this if you have <=9 bib refs
	\bibitem {rtai}{\it https://www.rtai.org/}
	\bibitem {sel4}{\it http://ssrg.nicta.com.au/projects/seL4/}
	\bibitem {linux-rt}{\it http://git.kernel.org/cgit/linux/kernel/git/rt/}
	\bibitem {xenomai}{\it http://xenomai.org/}
	\bibitem {adeos} Karim Yaghmour. Adaptive Domain Environment for Operating Systems.\\
	    \url{http://www.opersys.com/ftp/pub/Adeos/adeos.pdf}
	\bibitem {ChameleonRTOS}{\it http://www.xenomai.org/documentation/slides/Xenomai-OSMB-2007-01.pdf}
	\bibitem {posix-1003-1c}{\it http://standards.ieee.org/findstds/interps/1003-1c-95\_int/}
	\bibitem {xenomai-solo}{\it https://www.osadl.org/Xenomai-SOLO.xenomai-solo.0.html}
	\bibitem {x3-applications}{\it http://xenomai.org/running-applications-with-xenomai-3-x/}
	\bibitem {bbb}{\it http://beagleboard.org/BLACK}
	\bibitem {am335x}{\it http://www.ti.com/lsds/ti/processors/sitara/arm\_cortex-a8/am335x/overview.page}
	\bibitem {debian}{\it http://debian.beagleboard.org/images/bone-debian-7.5-2014-05-14-2gb.img.xz}
	\bibitem {kernel}{\it https://github.com/beagleboard/linux}
	\bibitem {p-rt}{\it https://www.kernel.org/pub/linux/kernel/projects/rt/3.14/}
	\bibitem {p-ipipe}{\it http://download.gna.org/adeos/patches/v3.x/arm/ipipe-core-3.14.39-arm-9.patch}
	\bibitem {git-x2.6}{\it http://git.xenomai.org/xenomai-2.6.git/}
	\bibitem {git-x3}{\it http://git.xenomai.org/xenomai-3.git/}
	\bibitem {stree}{\it http://people.seas.harvard.edu/~apw/stress/}
\end{thebibliography}
